---
title: "Paper: What Students Prefer to Do and What They Claim They Can Do Well: Interests Differentiation and Perceived Skills "
output: html_document
---

# ABOUT THE SCRIPT
- This a script containing code about the results section presented in the paper "What Students Prefer to Do and What They Claim They Can Do Well: Interests Differentiation and Perceived Skills" (submitted). Code used for data cleaning procedures is not presented in this script.

**18-REST2 items in Portuguese(BR) and English:**

-   r1 Operar máquinas para usinagem de peças./To operate machines.
-   r2 Fazer análises e experimentos em laboratórios./To perform data analyses and experiments in laboratories.
-   r3 Cantar em um coral./To sing in a choir.
-   r4 Estar disponível para ajudar as pessoas./To be available to help people.
-   r11 Negociar com clientes./To negotiate with clients.
-   r18 Arquivar documentos e notas importantes./To organize important documents and notes.
-   r7 Executar manutenções em máquinas e equipamentos./To perform maintanance in machines and other equipment.
-   r8 Explicar fenômenos físicos da natureza./To explain physics and nature phenomena.
-   r9 Apresentar números artísticos para uma plateia./To present artistic pieces to a crowd.
-   r16 Oferecer orientação às pessoas, grupos ou população sobre saúde e bem-estar./To offer well-being and health guidance for people, groups or the population. 
-   r17 Comercializar produtos e serviços./To sell products and services.
-   r22 Fazer a contabilidade de uma empresa./To do accounting in a business. 
-   r27 Projetar o sistema elétrico de uma casa./To think about the eletrical system of a house.
-   r14 Ler artigos e livros científicos./To read scientific papers and books.
-   r15 Participar da criação de cenários teatrais./To participate in the creation of theatre stages.
-   r24 Prestar serviços sociais em comunidades e bairros./To perform social services in communities.
-   r29 Convencer pessoas a comprar um produto./To convince people to buy a product.
-   r26 Inserir informações em uma base de dados./To insert information in a database.

```{r pacotes_db, echo= FALSE}
# LOAD PACKAGES -----------------------------------------------------------------
library(tidyverse)
library(psych)
library(holland)
library(stats)
library(lsr)
library(rstatix)
library(corrplot) 

```

## CALCULATING SCORES
```{r}
# LOAD DATA -----------------------------------------------------------------
# For example purposes, the dataset object will be named as dt_hol because the dataset is not available publicly 
# More information about the dataset can accessed on the paper
dt_hol <- readRDS("data/dt_hol.RDS")
```

- *RIASEC scores* were calculated as average scores. These scores are identified with the prefix holl_ + RIASEC letter.

- *Relative differentiation scores* were calculated using the following formula: Reference interest score -(average score of other five interests scores). These scores are identified with the prefix riasec_rlv OR the suffix _rlv3. 

```{r}
# AVERAGE RIASEC SCORES ------------------------------------------------------
# 18-REST2 MEASURE (Martins et al., 2024)
key_list <- list(holl_R = c("r1", "r7", "r27"),
                 holl_I = c("r2", "r8", "r14"),
                 holl_A = c("r3", "r9", "r15"),
                 holl_S = c("r4", "r16", "r24"),
                 holl_E = c("r11", "r17", "r29"),
                 holl_C = c("r18", "r22", "r26")
                 )

holl_scores <- psych::scoreItems(key_list, 
                          dt_hol, 
                          totals = FALSE,
                          missing=FALSE, 
                          min = 1, 
                          max = 5, 
                          digits = 2)

dt_hol <- cbind(dt_hol, holl_scores$scores)

```

```{r echo= FALSE, eval=FALSE}
# HOLLAND PACKAGE SCORES ------------------------------------------------------

# Overall Differentiation (dif7) and Elevation scores

# Add empty differentiation and elevation variables in the dt_hol  

dt_hol$holland_Difer7<-NA
dt_hol$holland_Elev <-NA

# Calculate scores based on the six RIASEC scores previously calculated 
# Columns 10 to 15 represent RIASEC scores

for (i in 1:nrow(dt_hol)) {
 dt_hol[i,16]<- dif_7_holland(as.numeric(dt_hol[i,10:15]),ind="DI7")
}

dt_hol$holland_Elev<-apply(dt_hol[,10:15],1, mean)

```

```{r echo= FALSE, eval=FALSE}
# SINGLE-INTEREST  PACKAGE SCORES ----------------------------------------------------------------

# Formula: Reference interest score -(average score of other five interests scores)
v_holl_r <- c("holl_I", "holl_A", "holl_S", "holl_E", "holl_C")
v_holl_i <- c("holl_R", "holl_A", "holl_S", "holl_E", "holl_C")
v_holl_a <- c("holl_R", "holl_I", "holl_S", "holl_E", "holl_C")
v_holl_s <- c("holl_R", "holl_I", "holl_A", "holl_E", "holl_C")
v_holl_e <- c("holl_R", "holl_I", "holl_A", "holl_S", "holl_C")
v_holl_c <- c("holl_R", "holl_I", "holl_A", "holl_S", "holl_E")

# R
dt_hol_final  <- dt_hol_final  %>% 
 mutate(holl_R_rlv = rowSums(.[, v_holl_r])
                      ) %>% 
 mutate(holl_R_rlv2 = holl_R_rlv/5) %>% 
 mutate(holl_R_rlv3 = holl_R - holl_R_rlv2)

# I
dt_hol_final  <- dt_hol_final  %>% 
 mutate(holl_I_rlv = rowSums(.[, v_holl_i])
                      ) %>% 
 mutate(holl_I_rlv2 = holl_I_rlv/5) %>% 
 mutate(holl_I_rlv3 = holl_I - holl_I_rlv2)

# A
dt_hol_final <- dt_hol_final  %>% 
 mutate(holl_A_rlv = rowSums(.[, v_holl_a])
                      ) %>% 
 mutate(holl_A_rlv2 = holl_A_rlv/5) %>% 
 mutate(holl_A_rlv3 = holl_A - holl_A_rlv2)

# S
dt_hol_final <- dt_hol_final  %>% 
 mutate(holl_S_rlv = rowSums(.[, v_holl_s])
                      ) %>% 
 mutate(holl_S_rlv2 = holl_S_rlv/5) %>% 
 mutate(holl_S_rlv3 = holl_S - holl_S_rlv2)

# E
dt_hol_final <- dt_hol_final  %>% 
 mutate(holl_E_rlv = rowSums(.[, v_holl_e])
                      ) %>% 
 mutate(holl_E_rlv2 = holl_E_rlv/5) %>% 
 mutate(holl_E_rlv3 = holl_E - holl_E_rlv2)

# C
dt_hol_final <- dt_hol_final  %>% 
 mutate(holl_C_rlv = rowSums(.[, v_holl_c])
                      ) %>% 
 mutate(holl_C_rlv2 = holl_C_rlv/5) %>% 
 mutate(holl_C_rlv3 = holl_C - holl_C_rlv2)

```

## RESULTS SECTION

### Interest elevation and overall differentiation across time 

#### Welch's t-tests - Gender groups
- Welch's t-test of interest elevation and overall differentiation comparing *gender groups*
- Comparisons were performed in the following order: F-M

```{r echo=FALSE, warning=FALSE, error=FALSE}
t.test(holland_Difer7 ~ gender, data = dt_hol)
cohensD(holland_Difer7 ~ gender, data = dt_hol) 

t.test(holland_Elev ~ gender, data = dt_hol)
cohensD(holland_Elev ~ gender, data = dt_hol) 

```

#### Welch's t-tests - School grades and gender groups
- Welch's t-test of interest elevation and overall differentiation comparing *school grades and gender groups*
- Comparisons were performed in the following order: F-M

```{r echo=FALSE, warning=FALSE, error=FALSE}
# SEPARATE DATASET USING SCHOOL YEAR 
dt_hol_final_5 <- dt_hol %>% 
 filter(serie == "5")

dt_hol_final_9 <- dt_hol %>% 
 filter(serie == "9")

dt_hol_final_12 <- dt_hol %>% 
 filter(serie == "12")

# T TESTS AND COHEN'S D FOR GENDER WITHIN EACH SCHOOL GRADE

#  5th grade
t.test(holland_Difer7 ~ gender, data = dt_hol_final_5)
cohensD(holland_Difer7 ~ gender, data = dt_hol_final_5) # 0.1475862
t.test(holland_Elev ~ gender, data = dt_hol_final_5)
cohensD(holland_Elev ~ gender, data = dt_hol_final_5) # 0.1500663

#  9th grade
t.test(holland_Difer7 ~ gender, data = dt_hol_final_9)
cohensD(holland_Difer7 ~ gender, data = dt_hol_final_9) # 0.1475862
t.test(holland_Elev ~ gender, data = dt_hol_final_9)
cohensD(holland_Elev ~ gender, data = dt_hol_final_9) # 0.1500663

#  12th grade
t.test(holland_Difer7 ~ gender, data = dt_hol_final_12)
cohensD(holland_Difer7 ~ gender, data = dt_hol_final_12) # 0.341822
t.test(holland_Elev ~ gender, data = dt_hol_final_12)
cohensD(holland_Elev ~ gender, data = dt_hol_final_12) # 0.04378867

```

#### ANOVA - School grandes in each gender group
- ANOVA  of interest elevation and overall differentiation comparing *school grades in each gender group*
```{r}
# SEPARATE DATASET USING GENDER GROUPS
dt_hol_final_f <- dt_hol %>% 
 filter(gender=="F")

dt_hol_final_m <- dt_hol %>% 
 filter(gender=="M")
```

```{r}
## INTEREST ELEVATION 

# ANOVA - Girls
dt_hol_final_f %>% 
anova_test(holland_Elev ~ as.factor(school_grade))
anova_summary(elev_girls_tab)

## Double check
elev_girls <- aov(holland_Elev  ~ as.factor(school_grade), dt_hol_final_f)
anova_summary(elev_girls)

# ANOVA - Boys
dt_hol_final_m %>% 
anova_test(holland_Elev ~ as.factor(school_grade))

## Double check
elev_boys <- aov(holland_Elev ~ as.factor(school_grade), dt_hol_final_m)
anova_summary(elev_boys)

```

### Single-interest differentiation across time

#### Welch's t-tests - School grades and gender groups
- Welch's t-test of single-interest differentiation comparing *gender groups in each school grades*
- Comparisons were performed in the following order: F-M

```{r}
#  5th grade
t.test(holl_R_rlv3 ~ gender, data = dt_hol_final_5)
cohensD(holl_R_rlv3 ~ gender, data = dt_hol_final_5) # 0.49

t.test(holl_I_rlv3 ~ gender, data = dt_hol_final_5)
cohensD(holl_I_rlv3 ~ gender, data = dt_hol_final_5) # 0.0798903

t.test(holl_A_rlv3 ~ gender, data = dt_hol_final_5)
cohensD(holl_A_rlv3 ~ gender, data = dt_hol_final_5) # 0.4791033

t.test(holl_S_rlv3 ~ gender, data = dt_hol_final_5)
cohensD(holl_S_rlv3 ~ gender, data = dt_hol_final_5) # 0.1180712

t.test(holl_E_rlv3 ~ gender, data = dt_hol_final_5)
cohensD(holl_E_rlv3 ~ gender, data = dt_hol_final_5) # 0.1156747

t.test(holl_C_rlv3 ~ gender, data = dt_hol_final_5)
cohensD(holl_C_rlv3 ~ gender, data = dt_hol_final_5) # 0.1375115

#  9th grade
t.test(holl_R_rlv3 ~ gender, data = dt_hol_final_9)
cohensD(holl_R_rlv3 ~ gender, data = dt_hol_final_9) # 0.91

t.test(holl_I_rlv3 ~ gender, data = dt_hol_final_9)
cohensD(holl_I_rlv3 ~ gender, data = dt_hol_final_9) # 0.18

t.test(holl_A_rlv3 ~ gender, data = dt_hol_final_9)
cohensD(holl_A_rlv3 ~ gender, data = dt_hol_final_9) # 0.54

t.test(holl_S_rlv3 ~ gender, data = dt_hol_final_9)
cohensD(holl_S_rlv3 ~ gender, data = dt_hol_final_9) # 0.39

t.test(holl_E_rlv3 ~ gender, data = dt_hol_final_9)
cohensD(holl_E_rlv3 ~ gender, data = dt_hol_final_9) # 0.19

t.test(holl_C_rlv3 ~ gender, data = dt_hol_final_9)
cohensD(holl_C_rlv3 ~ gender, data = dt_hol_final_9) # 0.10

#  12th grade
t.test(holl_R_rlv3 ~ gender, data = dt_hol_final_12)
cohensD(holl_R_rlv3 ~ gender, data = dt_hol_final_12) # 1.05

t.test(holl_I_rlv3 ~ gender, data = dt_hol_final_12)
cohensD(holl_I_rlv3 ~ gender, data = dt_hol_final_12) # 0.08

t.test(holl_A_rlv3 ~ gender, data = dt_hol_final_12)
cohensD(holl_A_rlv3 ~ gender, data = dt_hol_final_12) # 0.45

t.test(holl_S_rlv3 ~ gender, data = dt_hol_final_12)
cohensD(holl_S_rlv3 ~ gender, data = dt_hol_final_12) # 0.67

t.test(holl_E_rlv3 ~ gender, data = dt_hol_final_12)
cohensD(holl_E_rlv3 ~ gender, data = dt_hol_final_12) # 0.12

t.test(holl_C_rlv3 ~ gender, data = dt_hol_final_12)
cohensD(holl_C_rlv3 ~ gender, data = dt_hol_final_12) # 0.01

```

### Vocational interests versus social-emotional skills 

#### Correlations between RIASEC and social-emotional skills domains
```{r}
# RIASEC and SEMS vector with variables names

sems_riasec_names <- c("A", "C", "E", "N", "O",
                        "holl_R", "holl_I", "holl_A", "holl_S", "holl_E", "holl_C")
                        
# 5TH GRADE
sems_riasec_cor_5 <- cor(dt_hol_final_5[,sems_riasec_names], 
                             use="pairwise.complete.obs"
                        )

corrplot(sems_riasec_cor_5, method="color", 
         number.cex = 0.7, 
         addCoef.col = "black", 
         tl.col="black", tl.srt=45,
         diag=TRUE)


# 9TH GRADE
sems_riasec_cor_9 <- cor(dt_hol_final_5[,sems_riasec_names], 
                             use="pairwise.complete.obs"
                        )

corrplot(sems_riasec_cor_9, method="color", 
         number.cex = 0.7, 
         addCoef.col = "black", 
         tl.col="black", tl.srt=45,
         diag=TRUE)

# 12TH GRADE
sems_riasec_cor_12 <- cor(dt_hol_final_5[,sems_riasec_names], 
                             use="pairwise.complete.obs"
                        )

corrplot(sems_riasec_cor_12, method="color", 
         number.cex = 0.7, 
         addCoef.col = "black", 
         tl.col="black", tl.srt=45,
         diag=TRUE)
```

#### Congruence estimation

```{r}
### IMPORT CORRELATION TABLES FROM SOTO ET AL. (2021) IN R

### TABLES WERE NAMED AS RIASEC_SEMS_Soto_cong, RIASEC_SEMS_5th_cong (5th graders congruence), RIASEC_SEMS_9th_cong (9th graders congruence) and RIASEC_SEMS_12th_cong (12th graders congruence)

### CORRELATION MATRICES FOR RIASEC AND SEMS 
sems_riasec_cor_5_cong <- cor(
 dt_hol_final_5[,c(
  "O", 
  "C", 
  "E", 
  "A", 
  "N"
 )], 
 dt_hol_final_5[,c(
  "holl_R", 
  "holl_I", 
  "holl_A", 
  "holl_S", 
  "holl_E", 
  "holl_C"
 )], 
 use="pairwise.complete.obs"
)

sems_riasec_cor_9_cong <- cor(
 dt_hol_final_9[,c(
  "O", 
  "C", 
  "E", 
  "A", 
  "N"
 )], 
 dt_hol_final_9[,c(
  "holl_R", 
  "holl_I", 
  "holl_A", 
  "holl_S", 
  "holl_E", 
  "holl_C"
 )], 
 use="pairwise.complete.obs"
)

sems_riasec_cor_11_cong <- cor(
 dt_hol_final_11[,c(
  "O", 
  "C", 
  "E", 
  "A", 
  "N"
 )], 
 dt_hol_final_11[,c(
  "holl_R", 
  "holl_I", 
  "holl_A", 
  "holl_S", 
  "holl_E", 
  "holl_C"
 )], 
 use="pairwise.complete.obs"
)

### SEMS CONGRUENCE
cong_5th_Soto <- fa.congruence(sems_riasec_cor_5_cong, RIASEC_SEMS_Soto, structure = FALSE)

cong_9th_Soto <- fa.congruence(sems_riasec_cor_9_cong, RIASEC_SEMS_Soto, structure = FALSE)

cong_11th_Soto <- fa.congruence(sems_riasec_cor_11_cong, RIASEC_SEMS_Soto, structure = FALSE)

cong_5th_9th <- fa.congruence(sems_riasec_cor_5_cong, sems_riasec_cor_9_cong, structure = FALSE)

cong_5th_11th<- fa.congruence(sems_riasec_cor_5_cong, sems_riasec_cor_11_cong, structure = FALSE)

cong_9th_11th <- fa.congruence(sems_riasec_cor_9_cong, sems_riasec_cor_11_cong, structure = FALSE)


### OVERALL CONGRUENCE
cong_5th_Soto_g <- fa.congruence(RIASEC_SEMS_5th_cong, RIASEC_SEMS_Soto_cong, structure = FALSE)

cong_9th_Soto_g <- fa.congruence(RIASEC_SEMS_9th_cong, RIASEC_SEMS_Soto_cong, structure = FALSE)

cong_11th_Soto_g <- fa.congruence(RIASEC_SEMS_12th_cong, RIASEC_SEMS_Soto_cong, structure = FALSE)

cong_5th_9th_g <- fa.congruence(RIASEC_SEMS_5th_cong, RIASEC_SEMS_9th_cong, structure = FALSE)

cong_5th_11th_g <- fa.congruence(RIASEC_SEMS_5th_cong, RIASEC_SEMS_12th_cong, structure = FALSE)

cong_9th_11th_g <- fa.congruence(RIASEC_SEMS_9th_cong, RIASEC_SEMS_12th_cong, structure = FALSE)
```


### Interest elevation and overall differentiation and level of skills	 

#### Correlations between Interest elevation and Overall differentiation and social-emotional skills domains

```{r}
# Elevation, Diff and SEMS vector with variables names
                        
sems_elev_diff_names <- c("A", "C", "E", "N", "O",
                        "holland_Difer7", "holland_Elev")
                        
# 5TH GRADE
sems_elev_diff_cor_5 <- cor(dt_hol_final_5[,sems_elev_diff_names], 
                             use="pairwise.complete.obs"
                        )

corrplot(sems_elev_diff_cor_5, method="color", 
         number.cex = 0.7, 
         addCoef.col = "black", 
         tl.col="black", tl.srt=45,
         diag=TRUE)


# 9TH GRADE
sems_elev_diff_cor_9 <- cor(dt_hol_final_5[,sems_elev_diff_names], 
                             use="pairwise.complete.obs"
                        )

corrplot(sems_elev_diff_cor_9, method="color", 
         number.cex = 0.7, 
         addCoef.col = "black", 
         tl.col="black", tl.srt=45,
         diag=TRUE)

# 12TH GRADE
sems_elev_diff_cor_12 <- cor(dt_hol_final_5[,sems_elev_diff_names], 
                             use="pairwise.complete.obs"
                        )

corrplot(sems_elev_diff_cor_12, method="color", 
         number.cex = 0.7, 
         addCoef.col = "black", 
         tl.col="black", tl.srt=45,
         diag=TRUE)
```


### Single-interest-skill-differentiation hypothesis 

#### Correlations between skills domains and vocational interests 

```{r}
# RIASEC and SEMS vector with variables names

sems_riasec_names <- c("A", "C", "E", "N", "O",
                        "holl_R", "holl_I", "holl_A", "holl_S", "holl_E", "holl_C")
                        
# 5TH GRADE
sems_riasec_cor_5 <- cor(dt_hol_final_5[,sems_riasec_names], 
                             use="pairwise.complete.obs"
                        )

corrplot(sems_riasec_cor_5, method="color", 
         number.cex = 0.7, 
         addCoef.col = "black", 
         tl.col="black", tl.srt=45,
         diag=TRUE)


# 9TH GRADE
sems_riasec_cor_9 <- cor(dt_hol_final_5[,sems_riasec_names], 
                             use="pairwise.complete.obs"
                        )

corrplot(sems_riasec_cor_9, method="color", 
         number.cex = 0.7, 
         addCoef.col = "black", 
         tl.col="black", tl.srt=45,
         diag=TRUE)

# 12TH GRADE
sems_riasec_cor_12 <- cor(dt_hol_final_5[,sems_riasec_names], 
                             use="pairwise.complete.obs"
                        )

corrplot(sems_riasec_cor_12, method="color", 
         number.cex = 0.7, 
         addCoef.col = "black", 
         tl.col="black", tl.srt=45,
         diag=TRUE)
```

#### Correlations between skills domains and single-differentiated interests 

```{r}
sems_relatv_riasec_names <- c("A", "C", "E", "N", "O",
                        "holl_R_rlv3", "holl_I_rlv3", "holl_A_rlv3", "holl_S_rlv3",
                        "holl_E_rlv3", "holl_C_rlv3")

# 5TH GRADE
sems_relatv_cor_5 <- cor(dt_hol_final_5[,sems_relatv_riasec_names], 
                             use="pairwise.complete.obs"
                        )

corrplot(sems_relatv_cor_5, method="color", 
         number.cex = 0.7, 
         addCoef.col = "black", 
         tl.col="black", tl.srt=45,
         diag=TRUE)


# 9TH GRADE
sems_relatv_cor_9 <- cor(dt_hol_final_5[,sems_relatv_riasec_names], 
                             use="pairwise.complete.obs"
                        )

corrplot(sems_relatv_cor_9, method="color", 
         number.cex = 0.7, 
         addCoef.col = "black", 
         tl.col="black", tl.srt=45,
         diag=TRUE)

# 12TH GRADE
sems_relatv_cor_12 <- cor(dt_hol_final_5[,sems_relatv_riasec_names], 
                             use="pairwise.complete.obs"
                        )

corrplot(sems_relatv_cor_12, method="color", 
         number.cex = 0.7, 
         addCoef.col = "black", 
         tl.col="black", tl.srt=45,
         diag=TRUE)
```

### Circularity test 

- Figure 1 - Hexagons representations by school grade
* Domain level
* Pearson correlations
* N ~ 55578

#### 5th grade
```{r include= FALSE, echo=FALSE, warning=FALSE, message=FALSE}
# Create vector with RIASEC variables names
riasec_names <- c("holl_R", "holl_I", "holl_A", "holl_S", "holl_E", "holl_C"
                        )

riasec_cor_5 <- cor(dt_hol_final_5[,riasec_names], 
                             use="pairwise.complete.obs"
                        )

# Compute circularity test
riasec_circ_5 <- Circ_emp(55578, riasec_cor_5, konstrukt=c("R","I","A","S","E","C"), 
          showOutput=TRUE, Mplus_command="Mplus")

riasec_circ_5
```

#### 9th grade
```{r include= FALSE, echo=FALSE, warning=FALSE, message=FALSE}
# Create vector with RIASEC variables names
riasec_names <- c("holl_R", "holl_I", "holl_A", "holl_S", "holl_E", "holl_C"
                        )

riasec_cor_9 <- cor(dt_hol_final_9[,riasec_names], 
                             use="pairwise.complete.obs"
                        )

# Compute circularity test
riasec_circ_9 <- Circ_emp(82019, riasec_cor_9, konstrukt=c("R","I","A","S","E","C"), 
          showOutput=TRUE, Mplus_command="Mplus")

riasec_circ_9
```

#### 12th grade
```{r include= FALSE, echo=FALSE, warning=FALSE, message=FALSE}
# Create vector with RIASEC variables names
riasec_names <- c("holl_R", "holl_I", "holl_A", "holl_S", "holl_E", "holl_C"
                        )

riasec_cor_12 <- cor(dt_hol_final_12[,riasec_names], 
                             use="pairwise.complete.obs"
                        )

# Compute circularity test
riasec_circ_12 <- Circ_emp(70874, riasec_cor_12, konstrukt=c("R","I","A","S","E","C"), 
          showOutput=TRUE, Mplus_command="Mplus")

riasec_circ_12
```

